---
title: "R Notebook"
output: html_notebook
---

*Plot 1: Combined Histogram and Density Plot*

```{r}
# Load necessary libraries
library(ggplot2)
library(readr)
data <- read_csv("C:/Users/zubai/OneDrive/Desktop/Columbia/Assignments/EDAV/Project/cleaned_filtered.csv")

# Plot histogram with density overlay and improved theme
ggplot(data, aes(x = `International migrant stock: Both sexes (% total population)`)) +
  geom_histogram(aes(y = ..density..), bins = 30, fill = "#69b3a2", color = "black", alpha = 0.7) +
  geom_density(color = "#404080", size = 1.2, adjust = 1.5) +
  labs(
    title = "Histogram of International Migrant Stock (Both Sexes) \nPercentages",
    subtitle = "Density overlay emphasizes data distribution",
    x = "Migrant Stock (% of Population)", 
    y = "Density"
  ) +
  theme_minimal(base_size = 14) +  # Use a slightly larger base font size
  theme(
    plot.title = element_text(face = "bold", size = 16, hjust = 0.5, color = "#333333"),
    plot.subtitle = element_text(size = 12, hjust = 0.5, color = "#666666"),
    axis.title.x = element_text(face = "bold", color = "#444444"),
    axis.title.y = element_text(face = "bold", color = "#444444"),
    axis.text = element_text(size = 12, color = "#555555"),
    panel.grid.major = element_line(color = "gray80", size = 0.5),
    panel.grid.minor = element_blank(),  # Remove minor grid lines for clarity
    panel.background = element_rect(fill = "#f5f5f5", color = NA),
    plot.background = element_rect(fill = "#ffffff", color = NA)
  )


```

This histogram shows the distribution of international migrant stock as a percentage of the total population. The data is heavily *right-skewed*, with most countries having a migrant stock below 10% of their population. A few countries have exceptionally high percentages, nearing 100%, which are likely outliers. The density overlay highlights the concentration of values at the lower end, emphasizing that migration is relatively low as a proportion of the population in most countries.

*Plot 2: QQ Plot*

```{r}

# Load necessary libraries
library(ggplot2)
library(readr)

# Read the data
data <- read_csv("C:/Users/zubai/OneDrive/Desktop/Columbia/Assignments/EDAV/Project/cleaned_filtered.csv")

# Create Q-Q plot
ggplot(data, aes(sample = `International migrant stock: Both sexes (% total population)`)) +
  stat_qq(color = "blue", size = 1) +
  stat_qq_line(color = "red", linetype = "dashed", size = 1) +
  labs(
    title = "Q-Q Plot of International Migrant Stock",
    subtitle = "Comparing data distribution to a normal distribution",
    x = "Theoretical Quantiles",
    y = "Sample Quantiles"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold", size = 16, hjust = 0.5, color = "#333333"),
    plot.subtitle = element_text(size = 12, hjust = 0.5, color = "#666666"),
    axis.title = element_text(face = "bold", color = "#444444"),
    axis.text = element_text(size = 12, color = "#555555"),
    panel.grid.major = element_line(color = "gray80", size = 0.5),
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "#f5f5f5", color = NA),
    plot.background = element_rect(fill = "#ffffff", color = NA)
  )

```

The Q-Q plot compares the distribution of international migrant stock to a theoretical normal distribution. The data deviates significantly from the red diagonal line, particularly in the upper quantiles, confirming a right-skewed distribution. This suggests that while most countries have relatively low migrant stock, a small number of countries with extremely high migrant populations skew the distribution.

```{r}
# Load necessary libraries
library(readr)

# Read the data
data <- read_csv("C:/Users/zubai/OneDrive/Desktop/Columbia/Assignments/EDAV/Project/cleaned_filtered.csv")

# Remove NA values from the column
migrant_stock <- na.omit(data$`International migrant stock: Both sexes (% total population)`)

# Perform Shapiro-Wilk test
shapiro_test <- shapiro.test(migrant_stock)

# Print the test result
print(shapiro_test)

```

*Plot 3: Scatter plot*

```{r}
# Load necessary libraries
library(ggplot2)
library(readr)
library(scales)  # For better axis formatting

# Load the dataset
data <- read_csv("C:/Users/zubai/OneDrive/Desktop/Columbia/Assignments/EDAV/Project/cleaned_filtered.csv")

# Convert columns to numeric (remove commas if present)
data$`International migrant stock: Both sexes (number)` <- as.numeric(gsub(",", "", data$`International migrant stock: Both sexes (number)`))
data$`Total refugees and people in refugee-like situations (number)` <- as.numeric(gsub(",", "", data$`Total refugees and people in refugee-like situations (number)`))

# Remove outliers based on quantiles
q1_migrant <- quantile(data$`International migrant stock: Both sexes (number)`, 0.99, na.rm = TRUE)
q1_refugees <- quantile(data$`Total refugees and people in refugee-like situations (number)`, 0.99, na.rm = TRUE)

filtered_data <- data[
  data$`International migrant stock: Both sexes (number)` <= q1_migrant &
  data$`Total refugees and people in refugee-like situations (number)` <= q1_refugees, 
]

# Create the scatterplot with improved scales
ggplot(filtered_data, aes(x = `International migrant stock: Both sexes (number)`, 
                          y = `Total refugees and people in refugee-like situations (number)`)) +
  geom_point(alpha = 0.3, color = "#69b3a2") +  # Add points with transparency
  scale_x_continuous(labels = comma) +  # Format x-axis with commas
  scale_y_continuous(labels = comma) +  # Format y-axis with commas
  labs(
    title = "Scatterplot: International Migrant Stock vs Total Refugee \nPopulation",
    x = "International Migrant Stock (Number)",
    y = "Total Refugee Population (Number)"
  ) +
  theme_minimal(base_size = 14) +  # Minimal theme with larger font size
  theme(
    plot.title = element_text(face = "bold", size = 16, hjust = 0.5),
    axis.title.x = element_text(face = "bold"),
    axis.title.y = element_text(face = "bold"),
    panel.grid.major = element_line(color = "gray80", size = 0.5),
    panel.grid.minor = element_blank()
  )

```

The scatterplot shows the relationship between international migrant stock (number) and total refugee population. Most data points are clustered near the origin, indicating that many countries have low values for both metrics. However, a few outliers with extremely high migrant stock or refugee populations are visible, suggesting that only a handful of countries host a disproportionately large number of migrants and refugees. The lack of a clear trend suggests no strong linear correlation between these two variables.

*Plot 4 - Par Coord*

```{r}
# Load necessary libraries
library(GGally)
library(dplyr)
library(readr)

# Load the dataset
data <- read_csv("C:/Users/zubai/OneDrive/Desktop/Columbia/Assignments/EDAV/Project/cleaned_filtered.csv")

# Convert all character columns to UTF-8 encoding
data <- data %>%
  mutate(across(where(is.character), ~ iconv(., from = "", to = "UTF-8", sub = "")))

# Remove the row with "Total, all countries or areas"
data <- data %>% filter(`Region/Country/Area` != "Total, all countries or areas")

# Rename columns for simplicity
colnames(data)[which(colnames(data) == "International migrant stock: Both sexes (% total population)")] <- "Migrant_Stock"
colnames(data)[which(colnames(data) == "Total refugees and people in refugee-like situations (number)")] <- "Total_Refugees"
colnames(data)[which(colnames(data) == "Asylum seekers, including pending cases (number)")] <- "Asylum_Seekers"

# Convert numeric columns (remove commas if present)
data$Migrant_Stock <- as.numeric(data$Migrant_Stock)
data$Total_Refugees <- as.numeric(gsub(",", "", data$Total_Refugees))
data$Asylum_Seekers <- as.numeric(gsub(",", "", data$Asylum_Seekers))

# Filter out rows with missing values
data <- data %>%
  filter(!is.na(Migrant_Stock), !is.na(Total_Refugees), !is.na(Asylum_Seekers))

# Group by Region/Country/Area and calculate sums
grouped_data <- data %>%
  group_by(`Region/Country/Area`) %>%
  summarise(
    Migrant_Stock = sum(Migrant_Stock, na.rm = TRUE),
    Total_Refugees = sum(Total_Refugees, na.rm = TRUE),
    Asylum_Seekers = sum(Asylum_Seekers, na.rm = TRUE)
  )

# Select the top 10 countries based on Asylum Seekers
top_10_countries <- grouped_data %>%
  arrange(desc(Asylum_Seekers)) %>%
  slice(1:10)

# Normalize the data for better visualization
normalized_data <- top_10_countries %>%
  mutate(across(where(is.numeric), ~ (. - min(.)) / (max(.) - min(.))))

# Create a parallel coordinates plot
ggparcoord(
  data = normalized_data,
  columns = 2:4,  # Include Migrant Stock, Total Refugees, Asylum Seekers
  groupColumn = 1,  # Use Region/Country/Area as the grouping variable
  scale = "globalminmax", 
  alphaLines = 0.8,
  showPoints = TRUE
) +
  labs(
    title = "Parallel Coordinates Plot: Top 10 Regions",
    x = "Variables",
    y = "Normalized Values"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold", size = 16, hjust = 0.5),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

```

This parallel coordinates plot visualizes normalized values of migration metrics (Migrant Stock, Total Refugees, and Asylum Seekers) for the top 10 countries. The lines show significant variability across metrics for different countries. For example, some regions like "Northern America" and "United States of America" have high migrant stock but relatively lower refugee populations. Conversely, regions like "Africa" and "Asia" show higher refugee numbers but lower migrant stocks. This highlights the differing roles countries play in migration dynamics—some as destinations for migrants and others as hosts for refugees.

*Plot 5 - Heatmap*

```{r}
# Load necessary libraries
library(ggplot2)
library(tidyr)
library(dplyr)

# Use the top_10_countries dataset from the previous analysis
# Reshape the data for heatmap visualization
heatmap_data <- normalized_data %>%
  pivot_longer(
    cols = Migrant_Stock:Asylum_Seekers, 
    names_to = "Metric", 
    values_to = "Value"
  )

# Create the heatmap
ggplot(heatmap_data, aes(x = `Region/Country/Area`, y = Metric, fill = Value)) +
  geom_tile(color = "white") + # Add white borders between tiles
  scale_fill_gradient(low = "#f7fbff", high = "#08306b", na.value = "gray80") + # Gradient color scale
  labs(
    title = "Heatmap of Migration Drivers (Top 10 Regions)",
    x = "Country/Region",
    y = "Migration Metrics",
    fill = "Value (Normalized)"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1), # Rotate x-axis labels for better readability
    plot.title = element_text(face = "bold", size = 16),
    legend.title = element_text(face = "bold"),
    legend.position = "right"
  )

```

The heatmap provides an overview of migration metrics across the top 10 countries/regions. Darker shades indicate higher values. Regions like "United States of America" and "Northern America" exhibit high values for international migrant stock, while regions such as "Africa" and "Asia" show higher refugee populations. The variability in asylum seekers across regions is also evident, with some regions hosting significantly more asylum seekers than others. This visualization emphasizes the uneven distribution of migration pressures globally. These observations highlight key trends in migration metrics across different countries and regions, providing insights into global migration patterns and their disparities.
