```{r}
# Load Required Libraries
library(ggplot2)
library(readr)

# Load the Data
data <- read_csv("/Users/snehamaurya/Desktop/Project-EDAV/cleaned_filtered.csv")

# Filter Relevant Columns and Remove Missing Values
data_filtered <- data[!is.na(data$`International migrant stock: Both sexes (number)`), 
                      c("Year", "International migrant stock: Both sexes (number)")]

# Convert 'Year' to Factor for Proper Grouping
data_filtered$Year <- as.factor(data_filtered$Year)

# Create the Boxplot with Median Line
ggplot(data_filtered, aes(x = Year, y = `International migrant stock: Both sexes (number)`)) +
  geom_boxplot(outlier.color = "red", outlier.shape = 16, fill = "lightblue", color = "black") +
  scale_y_log10() +  # Apply log transformation for better visualization
  stat_summary(
    fun = median, 
    geom = "line", 
    aes(group = 1),

color = "green", 
    linetype = "dashed", 
    size = 1
  ) +
  labs(
    title = "International Migrant Stock by Year",
    x = "Year",
    y = "Log of International Migrant Stock (Number)"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10)
  )

```
This boxplot visualizes the distribution of international migrant stock (both sexes, in numbers), as given in our dataset across different years, using a logarithmic scale for better clarity. The following insights can be drawn:

##Key findings:
1. The median values (horizontal line in each box) indicate relatively stable migration patterns over the years, with slight year-to-year variations.

2. The presence of red outlier points highlights countries with exceptionally high migrant stocks in specific years.

3. The dashed green line connecting the medians shows a consistent trend in central tendencies over time, suggesting no drastic shifts in global migration patterns.

```{r}
# Create the Scatter Plot with a professional color scheme
ggplot(data, aes(
  x = `International migrant stock: Male (% total Population)`,
  y = `International migrant stock: Female (% total Population)`
)) +
  geom_point(aes(color = as.factor(Year)), size = 2, alpha = 0.7) +  # Points colored by year
  scale_color_manual(
    values = c("2005" = "lightgreen", "2010" = "orange", "2015" = "powderblue", "2020" = "pink")  # Custom colors
  ) +
  labs(
    title = "Male vs. Female Migrant Stock Percentage",
    x = "Male Migrant Stock (%)",
    y = "Female Migrant Stock (%)",
    color = "Year"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10),
    legend.position = "right"
  ) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "black", size = 1)  # Add equality line


```

This scatter plot examines the relationship between male and female international migrant stocks as a percentage of the total population. It offers insight into the gender distribution of migration patterns across different years.

##Key Observations:

1. Positive Correlation:
The data shows a strong positive correlation between male and female migrant stocks, as indicated by the clustering of points along the dashed equality line (black).This suggests that countries with higher male migrant stocks tend to have correspondingly higher female migrant stocks.

2. Temporal Trends:
Points are color-coded by year (2005, 2010, 2015, and 2020), allowing for the visualization of migration patterns over time.While no major year-to-year shifts in the overall trend are evident, the distribution of points reflects subtle variations in migration behavior across years.

3. Outliers:
Some points deviate significantly from the equality line, representing countries where the proportion of male and female migrants is imbalanced. These outliers could reflect specific geopolitical or cultural factors influencing gendered migration patterns in those countries.
```{r}
# Filter top 10 regions by total population of concern
top_regions <- data_region[order(-data_region$TotalPopulationConcern), ][1:10, ]

# Cleveland Dot Plot with Top 10 Regions
ggplot(top_regions, aes(
  x = TotalPopulationConcern,
  y = reorder(Region, TotalPopulationConcern)  # Sort regions by population
)) +
  geom_point(color = "blue", size = 3) +  # Add points
  geom_text(aes(label = scales::comma(TotalPopulationConcern)), hjust = -0.2, size = 3) +  # Add labels for values
  labs(
    title = "Cleveland Dot Plot of Total Population of Concern by Top 10 Regions",
    x = "Total Population of Concern (Number)",
    y = "Region"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10)
  ) +
  scale_x_continuous(labels = scales::comma)  # Format X-axis with commas


```
```{r}
ggplot(top_regions, aes(
  x = TotalPopulationConcern,
  y = reorder(Region, TotalPopulationConcern)
)) +
  geom_bar(stat = "identity", fill = "blue", color = "black") +  
  geom_text(aes(label = scales::comma(TotalPopulationConcern)), hjust = -0.2, size = 3) +  # Adding labels
  labs(
    title = "Total Population of Concern by Top 10 Regions",
    x = "Total Population of Concern (Number) for UNHCR",
    y = "Region (Country)"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
    axis.title.x = element_text(size = 15, face = "bold"),  # Enhanced X-axis title
    axis.title.y = element_text(size = 15, face = "bold"),  # Enhanced Y-axis title
    axis.text.x = element_text(size = 10),  # X-axis text size
    axis.text.y = element_text(size = 8)   # Reduced Y-axis text size for readability
  ) +
  scale_x_continuous(
    labels = scales::comma,
    limits = c(0, max(top_regions$TotalPopulationConcern) * 1.1),  # Add 10% buffer for spacing
    expand = expansion(mult = c(0, 0.1))  # Adjust spacing on the axis
  )


```
This bar chart presents the total population of concern to the UNHCR across the top 10 regions, ranked by their respective values. The primary goal of this visualization is to identify and compare the contributions of different regions to the total population under UNHCR's purview. By focusing on the top regions, the chart highlights the most significant contributors, enabling targeted analyses and facilitating resource allocation discussions.

1. Reasons for the Visualisation: 
The chart provides a clear ranking of regions based on their total population of concern.
This focused view enables decision-makers to prioritize regions that have the largest populations requiring assistance or intervention.Unlike the Cleveland Dot Plot, the bar chart is visually more intuitive for comparing values at a glance.

2. Key Findings:
Asia and Africa dominate the chart, contributing the highest numbers, with Asia surpassing 57 million and Africa exceeding 49 million. Other regions, such as Colombia, the Syrian Arab Republic, and Europe, also show substantial populations of concern, emphasizing diverse geographical challenges.Total global figures further contextualize the scale of the UNHCR’s work, with all regions combined exceeding 141 million.

##Highlights:
1. Understanding Global Disparities: The chart highlights significant differences in populations of concern across regions, making it clear which areas need urgent attention.

2. Facilitating Policy and Resource Allocation: Identifying regions with the highest populations enables stakeholders to make informed decisions about prioritizing interventions and allocating resources effectively.

3. Enhancing Clarity and Interpretation: The inclusion of precise numerical labels and the organized ranking of regions make the chart easy to read, enabling quick and accurate comparisons.

```{r}


library(ggplot2)
library(sf)
library(dplyr)
library(rnaturalearth)
library(rnaturalearthdata)

# Load world shapefile
world <- ne_countries(scale = "medium", returnclass = "sf")


# Load your data
data <- read_csv("/Users/snehamaurya/Desktop/Project-EDAV/cleaned_filtered.csv")

data_clean <- data %>%
  select(`Region/Country/Area`, Year, `Asylum seekers, including pending cases (number)`) %>%
  rename(Country = `Region/Country/Area`, AsylumSeekers = `Asylum seekers, including pending cases (number)`) %>%
  filter(!is.na(AsylumSeekers)) %>%
  group_by(Country) %>%
  summarize(AsylumSeekers = sum(AsylumSeekers, na.rm = TRUE), .groups = "drop")

# Merge with shapefile
world_data <- world %>%
  left_join(data_clean, by = c("name" = "Country"))

# Create the map
ggplot(world_data) +
  geom_sf(aes(fill = AsylumSeekers), color = "black") +
  scale_fill_gradient(low = "blue", high = "red", na.value = "grey50", name = "Asylum Seekers") +
  labs(
    title = "Global Distribution of Asylum Seekers",
    subtitle = "Visualizing pending asylum cases across countries",
    caption = "Source: UNHCR Data"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
    plot.subtitle = element_text(size = 12, hjust = 0.5),
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 8)
  )




```
The map is a visually appealing and accurate representation of the global distribution of asylum seekers. The gradient color scheme from blue to red effectively highlights regions with varying magnitudes of asylum seekers. The map provides a clear and intuitive understanding of the data.

##Predicted Trends and Observations:

1. Regional Concentrations:
High concentrations of asylum seekers are observed in specific regions, such as North America and parts of Europe. This could be attributed to the political and economic stability in these regions, making them desirable destinations for asylum seekers.

2. Disparities in Distribution:
Some countries in Africa and the Middle East also show significant numbers of asylum seekers, likely due to ongoing conflicts or humanitarian crises in these regions.

3. Patterns of Migration:
The map suggests a trend where asylum seekers predominantly target developed nations or regions that are geographically closer to conflict zones, indicating proximity and socio-political stability as key factors influencing asylum-seeking trends.


```{r}
library(dplyr)

# Ensure dataset is loaded and contains the required column names
# Rename columns if necessary to match the code

# Perform Min-Max Normalization
data_ridgeline_plot <- data_ridgeline_plot %>%
  group_by(`Region/Country/Area`) %>%
  mutate(
    NormalizedRefugeeNumbers = (`Total refugees and people in refugee-like situations (number)` -
                                min(`Total refugees and people in refugee-like situations (number)`, na.rm = TRUE)) /
                               (max(`Total refugees and people in refugee-like situations (number)`, na.rm = TRUE) -
                                min(`Total refugees and people in refugee-like situations (number)`, na.rm = TRUE))
  )

# Filter for the top 10 regions with the highest maximum normalized refugee numbers
top_normalized_regions <- data_ridgeline_plot %>%
  group_by(`Region/Country/Area`) %>%
  summarise(MaxNormalized = max(NormalizedRefugeeNumbers, na.rm = TRUE)) %>%
  top_n(10, MaxNormalized) %>%
  pull(`Region/Country/Area`)

# Filter the main dataset for only the top regions
data_normalized <- data_ridgeline_plot %>%
  filter(`Region/Country/Area` %in% top_normalized_regions)
```


```{r}
# Filter out irrelevant entries such as "Total, all countries or areas"
data_normalized_filtered <- data_normalized %>%
  filter(`Region/Country/Area` != "Total, all countries or areas") %>%
  filter(!is.na(NormalizedRefugeeNumbers))  # Exclude rows with NA normalized values

# Plot ridgeline for filtered regions
ggplot(data_normalized_filtered, aes(
  x = NormalizedRefugeeNumbers,
  y = reorder(`Region/Country/Area`, NormalizedRefugeeNumbers),
  fill = stat(x)
)) +
  geom_density_ridges_gradient(scale = 3, rel_min_height = 0.01) +
  scale_fill_gradientn(
    colors = c("lightblue", "blue", "darkblue", "red"),
    name = "Normalized Density",
    breaks = c(0, 0.25, 0.5, 0.75, 1),
    labels = c("0%", "25%", "50%", "75%", "100%")
  ) +
  labs(
    title = "Represenattion Refugee Numbers by Regions",
    subtitle = "Visualizing the density of normalized refugee numbers across key regions",
    x = "Normalized Refugee Numbers (0 to 1)",
    y = "Regions"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5, size = 12),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10),
    legend.position = "right",
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 8, margin = margin(t = 4))
  )


```
###From the ridgeline plot of normalized refugee numbers by region, we can see:

1. Europe displays the highest overall density of normalized refugee numbers, indicating it hosts the largest refugee population relative to its total population.
2. Asia and Africa also have significant concentrations of normalized refugee numbers, though less than Europe.
3. Pakistan, Germany, and Lebanon show some of the most prominent individual peaks, suggesting they host particularly high numbers of refugees compared to their total populations.
4. Regions like Uganda and Iran (Islamic Republic of) have lower normalized refugee numbers, meaning their refugee populations are smaller relative to their total populations.

The visual effectively conveys the disparities in refugee distributions across different geographic regions. This insight can help inform targeted humanitarian aid and resettlement efforts to address the unique challenges faced by the countries hosting the largest shares of refugees globally. Overall, this plot provides a high-level snapshot of the relative scale of refugee populations around the world.
